\documentclass[11pt]{article}

\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{graphicx} % Required for inserting images

\geometry{a4paper,left=2cm,right=2cm,top=2cm,bottom=2cm}

\linespread{1.2}

\begin{document}

\thispagestyle{empty}

\begin{center}
    \Large\textbf{Quantum Physics 1, Mid-term Examinaton}

\vspace{1cm}
\normalsize
\fbox{\color{blue}\textsf{April 23rd, 2025, 200 points}}

\vspace{0.5cm}
\fbox{You can use the fact that $\int_{-\infty}^{\infty} dz \exp{[-az^2]} = \sqrt{\pi / a}, \quad  \text{Re}(a)>0$, and $\Gamma(1/2) = \sqrt{\pi}$.}
\end{center}

\textbf{Problem 1.} [50 points]

Let us consider a Gaussian wave-packet wavefunction in one-dimension ($p_0,x_0,d>0$ are numbers below).
\begin{equation*}
    \psi (x) = \frac{1}{(2 \pi d^2 ) ^{1/4}} \exp{\left[ \frac{ip_0 x}{\hbar} - \frac{(x-x_0)^2}{4d^2} \right].}
\end{equation*}

Next recall that the deviation of some operator $\hat{A}$ (with expectation value specified in terms of wavefunction $\psi$) is defined by
\begin{equation*}
    \Delta \hat{A} \big|_{\psi} = \sqrt{ \langle ( \hat{A} - \langle \hat{A} \rangle )^2 \rangle } = \sqrt{ \langle \hat{A}^2 \rangle - \langle \hat{A} \rangle^2 }.
\end{equation*}

\begin{enumerate}[leftmargin=3.5em,itemsep=-3pt]
    \item[(1)] [20 points] Now for the above wavefunction $\psi (x)$, compute $\langle \hat{x} \rangle$, $\langle \hat{p} \rangle$.
    \item[(2)] [20 points] Similary, compute $\langle \hat{x}^2 \rangle$, $\langle \hat{p}^2 \rangle$.
    \item[(3)] [10 points] Show that
    \begin{equation*}
        \Delta \hat{x} \big|_{\psi} \Delta \hat{p} \big|_{\psi} = \frac{\hbar}{2}.
    \end{equation*}
\end{enumerate}

\textbf{Problem 2.} [60 points]

\vspace{-4mm}
\begin{enumerate}[leftmargin=3.5em,itemsep=-3pt]
    \item[(1)] [30 points] An exponential of an operator $\hat{A}$ defined to be
    \begin{equation*}
        \exp[\hat{A}] \equiv \sum_{j=0}^{\infty} \frac{\hat{A}^j}{j!}.
    \end{equation*}
\end{enumerate} 

\noindent
Let us consider a wavefunction $\psi(x)$ which can be expanded into Taylor series. Find the action the following operator $\hat{L}$ on $\psi (x)$.
\begin{equation*}
    \hat{L} = \exp{\left[ ax \frac{d}{dx} \right]}, \quad a = \text{real.}
\end{equation*}

\begin{enumerate}[leftmargin=3.5em,itemsep=-3pt]
    \item[(2)] [10 points] Consider an operator $\hat{S} = \alpha \hat{x} + i \hat{p}$, where $\alpha$ is a certain positive constant with the dimension of momentum/length. Find the zero eigenvalue (normalized) eigenfunction of $\hat{S}$ in \textit{position space}.
    \item[(3)] [10 points] Find the zero eigenvalue (normalized) eigenfunction of $\hat{S}$ in \textit{momentum space}.
    \item[(4)] [10 points] Compare the results of (2) and (3) and make a comment.
\end{enumerate}

\textbf{Problem 3.} [50 points]

Consider an \underline{attractive} delta function potential $V(x) = -\gamma \delta(x)$ $(\gamma > 0 )$ in one-dimension. The particle mass is $m$ and let us look for the \textit{bound states}.
\vspace{-4mm}

\begin{enumerate}[leftmargin=3.5em,itemsep=-3pt]
    \item[(1)] [30 points] Find the energy eigenvalues and the \underline{normalized} eigenfunctions,
    \item[(2)] [20 points] Find the expectation value of \textit{kinetic energy} and \textit{potential energy} of the bound state wavefunctions.
\end{enumerate}

\newpage

\thispagestyle{empty}
\textbf{Problem 4.} [40 points]

We have a one-dimensional potential step:
\begin{equation*}
    V(x) = 0, \ x<0, \quad V(x) = V_0, \ x \geq 0, \quad V_0 >0.
\end{equation*}

The boundary condition is such that the particle is \underline{coming from the left}.

\vspace{-4mm}
\begin{enumerate}[leftmargin=3.5em,itemsep=-3pt]
    \item[(1)] [30 points] Find the energy eigenvalues and the eigenfunctions for the energy $E \geq V_0$. You \textbf{don't} have to normalize the eigenfunctions.
    \item[(2)] [10 points] Compute the probability current density everywhere in space. Is it continuous at $x=0$?
\end{enumerate}
\end{document}
